---
title: SPI-M Summary DD/MM/YYYY
subtitle: This is a placeholder subtitle
output: 
  powerpoint_presentation:
    reference_doc: DHSC_template.pptx

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("pacman")) install.packages("pacman,knitr")
if (!require("ggplot2")) install.packages("ggplot2")

pacman::p_load(knitr)
library(ggplot2)
  
```


## Vaccine coverage in England

:::::::::::::: {.columns}
::: {.column}
```{r slide1, echo=FALSE}
ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv)) + scale_colour_dhsc_d(aesthetics = c("fill", "colour"))
```
:::
::: {.column}
Normal body text goes here. This is a new sentence. This is the *first* kind of emphasis. This is the **second** kind of emphasis.

This is a new paragraph.
:::
::::::::::::::

## Vaccine coverage by region and age as of Day Month

:::::::::::::: {.columns}
::: {.column}
```{r slide2, echo=FALSE}
ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv))
```
:::
::: {.column}
Normal body text goes here. This is a new sentence. This is the *first* kind of emphasis. This is the **second** kind of emphasis.

This is a new paragraph.
:::
::::::::::::::

## Vaccine coverage by ethnicity

:::::::::::::: {.columns}
::: {.column}
```{r slide3, echo=FALSE}
ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv))
```
:::
::: {.column}
Normal body text goes here. This is a new sentence. This is the *first* kind of emphasis. This is the **second** kind of emphasis.

This is a new paragraph.
:::
::::::::::::::
